---
title: "p8105_hw1_cx2347"
author: "Chuyuan XU"
date: "2025-09-11"
output: html_document
---

```{r setup, message= FALSE}
library(moderndive)
library(tidyverse)

data("early_january_weather")
```

## Problem 1

The dataset "early_january_weather" collect the hourly weather data for LGA, JKF and EWR in the first half month of January 2023.
There are 15 important variables included in the dataset: 
  'origin' (Weather Station); 'year', 'month', 'day' and 'hour' measure the time of recording; 
  'humid' (relative humidity); 'temp' (Temperature in F); 'dewp' (dewpoint in F);
  'wind_dr' (wind direction in degrees); 'wind_speed' (wind speed in mph); 'wind_gust' (gust speed in mph); 'precip' (precipitation in inches); 
  'pressure' (sea level pressure in millibars); 'visib' (visibility in miles); 'time_hour' (date and hour of the recording as a POSIXct date);\
The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.\
The mean temperature is `r round(mean(early_january_weather$temp),2)` F.

```{r problem1.1}
# plot the scatter plot
ggplot(
  early_january_weather, 
  aes(
    x=time_hour, 
    y=temp, 
    color=humid
    )
  ) + 
geom_point()
```

The scatter plot forms several continuous peaks. The range of the temp is between `r round(min(pull(early_january_weather, temp)), 2)` and `r round(max(pull(early_january_weather, temp)),2)` Fahrenheit degrees. The range of the time_hour drawn in the scatter plot is between `r min(pull(early_january_weather, time_hour))` and `r max(pull(early_january_weather, time_hour))`. There is an overall ascending trend of temperature shown in the scatter plot. At the summit of each peak, the humidity is usually higher than that at the valley.

```{r problem1.2}
# save the plot to the directory
ggsave(
    "early_january_weather_plot.pdf", 
    height=4, 
    width=6
  )
```
## Problem 2

```{r problem2, warning=FALSE}
# create the table
df <- 
  tibble(
    vec.num=rnorm(10),
    vec.log=(vec.num>0),
    vec.char=c("a","b","c","d","e","f","g","h","i","k"),
    vec.fac=factor(c("low","low","low","mid","mid","mid","mid","high","high","high"))
  )

# get the means of each variable
x1 = mean(pull(df, vec.num))
x2 = mean(pull(df, vec.log))
x3 = mean(pull(df, vec.char))
x4 = mean(pull(df, vec.fac))
```

The mean of the numeric vector is `r round(x1, 2)`.\
The mean of the logical vector is `r round(x2, 2)`.\
The mean of the character vector is `r x3`.\
The mean of the factor vector is `r x4`.\
\
The mean function works for the vec.num (the random sample of size 10 from a standard Normal distribution) and vec.log (the logical vector indicating whether elements of the sample are greater than 0) but doesn't work for the vec.char (a character vector of length 10) and vec.fac (a factor vector of length 10, with 3 different factor “levels”).

```{r problem2.2, eval=FALSE}
# trials of converting these values to numeric
as.numeric(pull(df, vec.log))
as.numeric(pull(df, vec.char))
as.numeric(pull(df, vec.fac))
```
The logical vector and the factor vector can be converted to numeric values correctly by the as.numeric() function.
The character vector cannot be converted to numeric values correctly by the as.numeric() function.

This explains why the means() function works for the logical vector as they are already in a numeric format and why the means() function doesn't work for the character vector as we can't extract numeric values from characters. However, it still can't resolve why the means() function doesn't work for the factor vector. We can convert the factor vector into numeric values, but we can't perform the function means() on them.
